Another Look at Regularized Discriminant Analysis for High-Dimensional Classification

### 2013-02-14 reports/paper-rda.tex
Rewrote introduction. Significant changes to the discussion on RDA and our proposed generalization.

\newtheorem{cor}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{thm}{Theorem}

@@

Microarrays have enabled researchers to measure gene-expression levels of thousands of genes in a single experiment to identify a subset of genes and their expression patterns that either characterize a particular cell state or facilitate an accurate prediction of future cell state.  We focus on the classification of tissues and other biological samples into one of $K$ disjoint, predefined cancer types, where our goal is to determine accurately the class membership of a new tissue sample based on gene-expression microarray data. However, microarray data typically consist of a large number of gene expressions $p$ relative to the sample size $n$. Moreover, standard benchmark microarray data sets consist of between $p = 10,000$ and $55,000$ probes and fewer than $n = 100$ observations, in which case standard supervised-learning methods exhibit poor classification performance, due to the \emph{curse of dimensionality} \citep{Bellman:1961tn}. Furthermore, well-known classifiers, including linear discriminant analysis \emph{(LDA)} and quadratic discriminant analysis \emph{(QDA)}, are incalculable when $p > n$ because the sample covariance matrices employed in both classifiers are singular. As a result, many researchers, including \cite*{Merchante:2012vk}, \cite{Witten:2011kc}, \cite*{Pang:2009ik}, \cite*{Huang:2010ju}, and \cite*{Clemmensen:2011kr}, have restricted the \emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that must be estimated to ensure that the resulting classifiers are calculable.  For instance, \cite{Dudoit:2002ev} have proposed a modified \emph{LDA} classifier where the class covariance matrices are assumed to be diagonal.

Alternatively, several researchers, including \cite*{Zhang:2010va}, \cite*{Guo:2007te}, \cite{Srivastava:2007ud}, and \cite{Mkhadri:1995jp}, have proposed regularization methods to improve the estimation of the population covariance matrices before applying the \emph{LDA} and \emph{QDA} classifiers. Covariance-matrix regularization methods often are based on the well-known ridge-regression approach from \cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices to ensure that the estimator is positive definite. This shrinkage stabilizes eigenvalues that are near zero to ensure that the inverse of the the sample covariance matrix is well-posed. For an excellent overview of the regularization problem in discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.

Here, we consider specifically the  popular regularized discriminant analysis (\emph{RDA}) classifier from \cite{Friedman:1989tm} that incorporates a convex combination of the sample covariance matrices from the \emph{LDA} and \emph{QDA} classifiers. Because the resulting biased covariance-matrix estimator often remains singular in the small-sample, high-dimensional setting, \cite{Friedman:1989tm} applies covariance-matrix shrinkage, where the resulting covariance-matrix estimator is shrunken towards a scaled identity matrix. \cite{Friedman:1989tm} have shown that the \emph{RDA} classifier can attain improved classification accuracy, compared to the \emph{LDA} and \emph{QDA} classifiers when $n > p$.

In this paper we present a generalization of the \emph{RDA} classifier in terms of matrix functions. We show that several common covariance-matrix regularization methods can be specified in terms of a matrix function including the \emph{RDA} classifier as well as methods from \cite{Srivastava:2007ww} and \cite*{Thomaz:2006ef} and several sparse diagonal classifiers, including the diagonal \emph{LDA} (\emph{DLDA}) and \emph{QDA} (\emph{DQDA}) classifiers from \cite{Dudoit:2002ev}, and some of its variants including those proposed by \cite{Pang:2009ik}.  Our generalization elucidates some properties of these covariance-regularization methods and the relationships among them. Moreover, our generalization allows us to view the covariance-matrix regularization methods in terms of eigenvalue thresholds, which further allows us to identify situations where each method is ineffective and ineffective. From a perspective of eigenvalue thresholds, we propose two new classifiers that are often superior to the \emph{RDA} classifier in terms of classification performance because both methods employ a more robust eigenvalue thresholding to regularize the convex combination of covariance matrix estimators employed in the \emph{RDA} classifier. Our two proposed methods also have an advantage over the \emph{RDA} classifier because both methods employ dimension reduction by reducing the number of basis functions included in the \emph{RDA} classifier, thereby reducing the variance of the resulting classifiers.

Using three small-sample, high-dimensional microarray data sets, we demonstrate that our two proposed classifiers often yield superior classification accuracy to the \emph{RDA}, \emph{MDEB}, and \emph{NLDA} classifiers. We also include in our comparison the alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of diagonal discriminant analysis from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}. Additionally, we examine the classification accuracy of the classifiers considered using \textbf{three} simulation configurations. Furthermore, we show that our proposed eigenvalue-threshold methods are more numerically stable than the other considered regularization methods using the simulated configurations.

Quote from \cite{Ye:2009gd}: ``There is a lack of a systematic study to explore the commonalities and differences of these algorithms, as well as their intrinsic relation- ship. This has been a challenging task, since different algorithms apply completely different schemes when dealing with the singularity problem.'' Cite this as an issue with LDA and that regularization has been less explored with QDA. Our formulation provides the necessary tools to study this is in detail.

\cite{Ye:2006jm} have shown that shrinking the zero eigenvalues does not improve the classification performance of the \emph{RDA} classifier when $\lambda = 1$. Consequently, we instead shrink only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, which has two additional benefits. First, by considering only the nonzero eigenvalues of $\widehat{\bm\Sigma}_k(\lambda)$, we effectively ignore the eigenvectors corresponding to the zero eigenvalues, which reduces significantly the number of parameters that we incorporate into our \emph{RDA} classifier because $p \gg q_k$ if $p \gg N$. Furthermore, the size of the matrix $\bm V_k$ is $p \times N$ as opposed to a $p \times p$ matrix in the original definition of the \emph{RDA} classifier. For significantly large $p$, the reduced size of $\bm V_k$ expedites calculation of the \emph{RDA} classifier and reduces its memory usage.\footnote{TODO: Provide an example for a realistic $p$ and $N$.}

As noted by \cite{Hoyle:2011vt}, \cite{Raudys:1998dd}, and \cite{Ramey:2011ji}, the pseudoinverse is often inadvisable because the resulting classifier often yields a sharp decrease in classification accuracy if $p \approx n$.\footnote{TODO: Contrast the computational efficiency we gain with the need to stabilize the smallest eigenvalues (those near $q$).}

We prove several results using arguments similar to \cite{Ye:2006tq}.

@@

\section{Regularized Discriminant Analysis}

In discriminant analysis we wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$. Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a \times b$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate normal distribution with known mean vector  $\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite matrices. Because the parameters are typically unknown, we estimate the unknown parameters $\bm \mu_k$ and $\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the MLEs into the \emph{QDA} classifier. Assuming, for simplicity, that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ using the sample \emph{QDA} classifier:
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume further that the covariance matrix parameters are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then the pooled sample covariance matrix $\widehat{\bm\Sigma}$ is applied instead for each class, where
\begin{align}
	\widehat{\bm\Sigma} = \frac{1}{n} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
\end{align}
In this case \eqref{eq:qda} reduces to Fisher's linear discriminant analysis (LDA) classifier, and the log-determinant is omitted because it is constant across all classes.

@@

The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{\bm \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}, and this bias increases as $p$ increases relative to $n_k$ \citep{TODONeedCitation}. Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0. Therefore, although more feature information is available to discriminate among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does not exist.

Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$ so that \eqref{eq:qda} is calculable and has reduced variability in the standard bias-variance tradeoff. These methods often can be written concisely as a matrix function, which we define here using the notation of \cite{Izenman:2008gm}. Let $\bm A \in \mathbb{R}_{p \times p}$ be symmetric, and let $\phi:\mathbb{R} \rightarrow \mathbb{R}$ be monotonic. Then
\begin{align}
	\phi(\bm A) = \sum_{j = 1}^p \phi(e_j) \bm v_j \bm v_j'\label{eq:spectral-decomp}
\end{align}
is a matrix function defined explicitly in terms of the eigenvalues of $\bm A$. For this reason, $\phi$ is often called a spectral transform \citep{Zhu:2010vu} or a transfer function \citep{Ye:2009gd}. Special cases of \eqref{eq:spectral-decomp} include the spectral decomposition of $\bm A$ when $\phi(e_j) = e_j$, the matrix square-root of $\bm A$ when $\phi(e_j) = e_j^{1/2}$, and, if it exists, the inverse of $\bm A$ when $\phi(e_j) = e_j^{-1}$ \citep{Harville:2008wja}.

@@

Another common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
\begin{align}
	\phi(e_{jk}) = e_{jk} + \gamma_k, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression. Equation \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards the $p$-dimensional identity matrix $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$. Specifically, the zero eigenvalues are replaced with $\gamma_k$, resulting in $\widehat{\bm \Sigma}_k + \gamma_k \bm I_p$ being positive definite. For additional covariance-matrix regularization methods, see \cite{Ye:2009gd} and \cite{Ramey:2011ji}.

@@

Here, we define our generalization of the \emph{RDA} classifier by first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear combination of the crossproducts of the training observations centered by their respective class sample means. Then, we generalize the adjustment of the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ by introducing a more general matrix-function formulation. This formulation allows us to adjust the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ in a more general manner so that we are not limited to ridge-like estimators.
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda \widehat{\bm\Sigma}.\label{eq:sig-lambda-alternative}
\end{align}
as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we no longer have need for the convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i = 1, \ldots, N$, each centered by its class sample mean, we attain a clear interpretation of $\widehat{\bm\Sigma}_k(\lambda)$. We have that
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'} \nonumber \\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of each of the $N$ observations to estimating $\bm \Sigma_k$. That is, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$ observations. Below, we will show that the pooling operation is advantageous in increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma}_k)$ for $0 < \lambda \le 1$. Notice that if $\lambda  = 0$, then the observations from the remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to $\widehat{\bm\Sigma}$.

As we have discussed, several eigenvalue adjustment methods have been proposed to increase eigenvalues (approximately) equal to 0. Here, we define a more general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} by applying the matrix function $\phi_k(\widehat{\bm \Sigma}_k(\lambda))$. To simplify the notation, we write $\tilde{\bm \Sigma}_k = \phi_k(\widehat{\bm \Sigma}_k(\lambda))$. Thus, we have the generalized \emph{RDA} (\emph{GRDA}) classifier
\begin{align}
	D_{GRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k|, \label{eq:grda}
\end{align}

>>>>

\newtheorem{cor}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{thm}{Theorem}

@@

Microarrays have enabled researchers to measure gene-expression levels of thousands of genes in a single experiment to identify a subset of genes and their expression patterns that either characterize a particular cell state or facilitate an accurate prediction of future cell state.  We focus on the classification of tissues and other biological samples into one of $K$ disjoint, predefined cancer types, where our goal is to ascertain the class membership of a new tissue sample based on gene-expression microarray data. However, microarray data typically consist of a large number of gene expressions $p$ relative to the sample size $N$. Moreover, standard benchmark microarray data sets usually consist of between $p = 10,000$ and $55,000$ probes and fewer than $N = 100$ observations, in which case standard supervised-learning methods exhibit poor classification performance, due to the \emph{curse of dimensionality} \citep{Bellman:1961tn}. Furthermore, well-known classifiers, including linear discriminant analysis \emph{(LDA)} and quadratic discriminant analysis \emph{(QDA)}, are incalculable when $p > N$ because the sample covariance matrices employed in both classifiers are singular. As a result, many researchers, including \cite*{Merchante:2012vk}, \cite{Witten:2011kc}, \cite*{Pang:2009ik}, \cite*{Huang:2010ju}, and \cite*{Clemmensen:2011kr}, have introduced restrictions on the population covariance matrices in the \emph{LDA} and \emph{QDA} models to reduce number of covariance parameters that must be estimated to ensure that the resulting classifiers are calculable.  For instance, \cite{Dudoit:2002ev} have proposed the diagonal \emph{LDA} (\emph{DLDA}) classifier, a modification of the \emph{LDA} classifier where the class covariance matrices are assumed to be diagonal.

Alternatively, several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite{Srivastava:2007ud}, \cite{Ye:2006vx}, \cite*{Thomaz:2006ef}, \cite{Ye:2005uu}, \cite{Mkhadri:1995jp}, and \cite{Peck:1982tt}, have proposed regularization methods to improve the estimation of the population covariance matrices before applying the \emph{LDA} and \emph{QDA} classifiers. Covariance-matrix regularization methods often are based on the well-known ridge-regression approach from \cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices to ensure that the estimator is positive definite. Typically, a positive shrinkage value is added to each eigenvalue estimate, replacing the zero eigenvalues with the shrinkage constant and stabilizing the eigenvalues near zero to ensure that the inverse of the the sample covariance matrix is well-posed. This common form of covariance-matrix regularization is equivalent to summing the original covariance-matrix estimator with the identity matrix scaled by the shrinkage constant. For an excellent overview of the regularization problem in discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.

Here, we consider specifically the popular regularized discriminant analysis (\emph{RDA}) classifier from \cite{Friedman:1989tm}, who has proposed a weighted average of the covariance-matrix estimators from the \emph{LDA} and \emph{QDA} classifiers. Because the resulting biased covariance-matrix estimator can remain singular or nearly singular, \cite{Friedman:1989tm} has additionally incorporated applied covariance-matrix shrinkage, where the shrinkage constant is a function of the average eigenvalue of the original estimator. \cite{Friedman:1989tm} has shown that the \emph{RDA} classifier can attain improved classification accuracy, compared to the \emph{LDA} and \emph{QDA} classifiers when $p > N$. However, in its original form, the \emph{RDA} classifier is impractical when $p \gg N$ because the \emph{RDA} decision rule requires the calculation of the inverses and determinants of the $p \times p$ covariance-matrix estimators for each class. Furthermore, the \emph{RDA} classifier has two tuning parameters that are selected via cross-validation, which adds a substantial computational burden to the model-selection process. Despite the introduction of a computational shortcut by \cite{Friedman:1989tm} for model selection to alleviate some of the computation, the \emph{RDA} classifier is ill-advised on modern data sets because the model selection requires on the order of $Kp^3$ calculations for each tuning-parameter pair for each fold in the cross-validation scheme.

In this paper we address the computational shortcomings of the \emph{RDA} classifier. Applying a technique similar to \cite{Ye:2006tq}, we decompose the \emph{RDA} classifier and show that the sum related to the null space of the pooled sample covariance matrix can be discarded. In turn, this omission provides a substantial increase in the computational efficiency of the \emph{RDA} classifier and its model-selection. In fact, we show rigorously that the inverse and determinants can be reduced to calculations of $n_k \times n_k$ matrices, where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$. This reduction in calculation is clearly advantageous because $n_k$ is often quite small, say, 15--20, in microarray data sets. Thus, we demonstrate that the calculation of the \emph{RDA} decision rule involving inverses and determinants of extremely large covariance matrices is equivalent a decision rule involving inverses and determinants on much smaller matrices.

Our improvements to the calculation and model selection of the \emph{RDA} classifier yields a natural generalization of the \emph{RDA} classifier. In our proposed generalization we first simplify the pooling matrix in the \emph{RDA} classifier by constructing instead a convex combination of the covariance-matrix estimators employed in the \emph{LDA} and \emph{QDA} classifiers. This simplification is advantageous in that it improves the interpretation of the pooling operation of the \emph{RDA} classifier. Next, we consider covariance-matrix regularization in terms of matrix functions, which includes a large family of shrinkage methods, such as the methods from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and a wide variety of other shrinkage methods studied by \cite{Ramey:2011ji} and \cite*{Xu:2009fl}. While a large number of these methods can be viewed as variants of ridge regression, we consider the more general case, where a real monotonic function is applied to the eigenvalue estimates. We show that adjusting the zero eigenvalues with a general matrix function has no effect on the \emph{RDA} classifier. That is, we show that the \emph{RDA} classifier is invariant to any adjustments of the eigenvalues corresponding to the null space of the pooled sample covariance matrix. Moreover, there is no classificatory information available in the null space of the pooled sample covariance matrix.  Hence, no gain in classification performance can be obtained by adjusting these zero eigenvalues. Although \cite{Ji:2008wp} and \cite{Ye:2006jm} have provided a similar invariance property for the \emph{LDA} decision rule, the \emph{LDA} classifier involves the assumption that the population covariance matrices are equal, which is often too stringent \citep{Clemmensen:2011kr}. The \emph{RDA} classifier allows us to relax this assumption by including unrestricted covariance matrices in the model. The \emph{RDA} classifier is then a compromise between the two approaches, favoring the stronger linear assumption if a larger value of the pooling parameter is selected as optimal. Additionally, in the linear case  \cite{Ye:2009gd} have remarked that  a lack of a systematic study has been conduced to explore the relationships among various shrinkage methods. Our formulation provides a general framework from which we can elucidate properties of covariance-regularization methods and the relationships among them in the linear case, the quadratic case, and the cases in between.

In addition to the improvement in computational performance of classifiers with regularized covariance matrices, we demonstrate that our generalized \emph{RDA} classifier is competitive and often superior to several recent classifiers designed for small-sample, high-dimensional data. In particular, we compare our generalized RDA classifier with an alternatively-defined \emph{RDA} classifier from \cite*{Hastie:2008dt}, the regularized \emph{LDA} classifier from \cite{Guo:2007te}, a recently proposed variant of the \emph{DLDA} classifier from \cite*{Tong:2012hw}, and the penalized \emph{LDA} classifier from \cite{Witten:2011kc}. We conduct our comparison of the classification accuracy of the competing classifiers using three small-sample, high-dimensional microarray data sets and examine the classification accuracy of the competing classifiers using \textbf{three} simulation configurations.

@@

\section{Regularized Discriminant Analysis}
\label{sec:rda}

In\footnote{TODO: Add paragraph of other notation here, including null space, orthogonal complement, and direct sum. Example: $\bm A \oplus \bm B$ denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in \mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}.} discriminant analysis we wish to assign correctly an unlabeled $p$-dimensional observation vector $\bm x$ to one of $K$ unique, known classes (or populations) by constructing a classifier from $n$ training observations that can accurately predict the class membership of $\bm x$. Let $\bm x_i = (x_{i1}, x_{i2}, \ldots, x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots, n)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$, where $\mathbb{R}_{a \times b}$ denotes the matrix space of all $a \times b$ matrices over the real field $\mathbb{R}$. We assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density function (PDF) of the $k$th class, $p(\omega_k)$ is prior probability of class membership of the $k$th class.

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate normal distribution with known mean vector  $\bm\mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where $\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$ positive-definite matrices. Because the parameters are typically unknown, we estimate the unknown parameters $\bm \mu_k$ and $\bm\Sigma_k$ with their maximum likelihood estimators (MLEs) and substitute the MLEs into the \emph{QDA} classifier. Assuming that the prior probabilities of class membership $p(\omega_k)$ are equal for $k = 1, \ldots, K$, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the sample \emph{QDA} classifier
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\widehat{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\widehat{\bm\Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm\Sigma}_k$ are the MLEs for $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume further that the population covariance matrices are equal for each class (i.e., $\bm\Sigma_k = \bm\Sigma$ for all $k$), then the pooled sample covariance matrix $\widehat{\bm\Sigma}$ is substituted for $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\bm\Sigma} = \frac{1}{n} \sum_{k=1}^K n_k \widehat{\bm\Sigma}_k. \label{eq:pooled-cov}
\end{align}
is the MLE for $\bm \Sigma$. In this case \eqref{eq:qda} reduces to Fisher's \emph{LDA} classifier, and the log-determinant is omitted because it is constant across all classes.

@@

The smallest eigenvalues and the directions associated with their eigenvectors highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{\bm \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are underestimated \citep{Seber:2004uh}, Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p - n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are 0. Hence, the bias of the eigenvalue estimators increases as $p$ increases relative to $n_k$. Furthermore, although more feature information is available to discriminate among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does not exist.

Several regularization methods, such as the methods considered by \citet*{Guo:2007te}, \cite{Mkhadri:1995jp}, and \citet*{Xu:2009fl}, have been proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$ so that \eqref{eq:qda} is calculable and has reduced variability in the standard bias-variance tradeoff. These methods often can be written concisely as a matrix function, which we define here using the notation of \cite{Izenman:2008gm}. Let $\bm A \in \mathbb{R}_{p \times p}$ be symmetric, and let $\phi:\mathbb{R} \rightarrow \mathbb{R}$ be monotonic. Then
\begin{align}
	\phi(\bm A) = \sum_{j = 1}^p \phi(e_j) \bm v_j \bm v_j'\label{eq:spectral-decomp}
\end{align}
is a matrix function defined explicitly in terms of the eigenvalues of $\bm A$. For this reason, $\phi$ is often called a spectral transform \citep*{Zhu:2010vu} or a transfer function \citep{Ye:2009gd}. Special cases of \eqref{eq:spectral-decomp} include the spectral decomposition of $\bm A$ when $\phi(e_j) = e_j$, the matrix square-root of $\bm A$ when $\phi(e_j) = e_j^{1/2}$, and, if it exists, the inverse of $\bm A$ when $\phi(e_j) = e_j^{-1}$ \citep{Harville:2008wja}.

@@

Another common form of the covariance-matrix regularization applies a shrinkage factor $\gamma_k > 0$, so that
\begin{align}
	\phi(e_{jk}) = e_{jk} + \gamma_k, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression. Equation \eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance matrix $\widehat{\bm\Sigma}_k$ towards the $p$-dimensional identity matrix $\bm I_p$, thereby increasing the eigenvalues of $\widehat{\bm\Sigma}_k$ by $\gamma_k$. Specifically, the zero eigenvalues are replaced with $\gamma_k$, resulting in $\widehat{\bm \Sigma}_k + \gamma_k \bm I_p$ being positive definite. For additional covariance-matrix regularization methods, see \cite{Ramey:2011ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

Here, we define our generalization of the \emph{RDA} classifier by first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in \eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear combination of the crossproducts of the training observations centered by their respective class sample means. Then, we generalize the adjustment of the eigenvalues of $\widehat{\bm \Sigma}_k(\lambda)$ by introducing a more general matrix-function formulation.  We define
\begin{align}
  \widehat{\bm\Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm\Sigma}_k + \lambda \widehat{\bm\Sigma}.\label{eq:sig-lambda-alternative}
\end{align}
as a convex combination of the MLEs of $\bm \Sigma_k$ and $\bm \Sigma$, where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we no longer have need for the convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i = 1, \ldots, N$, each centered by its class sample mean, we attain a clear interpretation of $\widehat{\bm\Sigma}_k(\lambda)$. We have that
\begin{align}
	\widehat{\bm\Sigma}_k(\lambda) &= \widehat{\bm\Sigma}_k + \lambda (\widehat{\bm\Sigma} - \widehat{\bm\Sigma}_k) \nonumber\\
%	&= \widehat{\bm\Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm\Sigma}_{k'}  - \widehat{\bm\Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm\Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm\Sigma}_{k'} \nonumber \\
%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) \bm x_i \bm x_i' +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) \bm x_i \bm x_i' \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i',\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of each of the $N$ observations to estimating $\bm \Sigma_k$. That is, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$ observations. Below, we will show that the pooling operation is advantageous in increasing the rank of each $\widehat{\bm\Sigma}_k(\lambda)$ from rank$(\widehat{\bm\Sigma}_k)$ to rank$(\widehat{\bm\Sigma}_k)$ for $0 < \lambda \le 1$. Notice that if $\lambda  = 0$, then the observations from the remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to $\widehat{\bm\Sigma}$.

As we have discussed, several eigenvalue adjustment methods have been proposed to increase eigenvalues (approximately) equal to 0. In this section we define a more general eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} by applying the matrix function $\phi_k(\widehat{\bm \Sigma}_k(\lambda))$. To simplify the notation, we write $\tilde{\bm \Sigma}_k = \phi_k(\widehat{\bm \Sigma}_k(\lambda))$. Thus, we have the generalized \emph{RDA} (\emph{GRDA}) classifier
\begin{align}
	D_{GRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)'\tilde{\bm\Sigma}_k^{-1}(\bm x - \xbar_k)  + \ln |\tilde{\bm\Sigma}_k|, \label{eq:grda}
\end{align}
The specification of $\phi_k$ in \eqref{eq:grda} allows for a large family of eigenvalue shrinkage methods to be examined and for their relationships to be examined in depth.
